---
title: "Generating questions and solutions documents from the same R Markdown document"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating questions and solutions documents from the same R Markdown document}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This package provides a simple extra helper function when knitting an `.Rmd` file in which we have questions and solutions.

We ensure elements which are questions are either included as Rmarkdown text or as code chunk with `include=TRUE`.

We suppress or show the solutions with the knitr global document `include` chunk option controlled by `knitr::opts_chunk$set(include)`. And paragraph Rmarkdown text for the solutions is included in an `asis` code chunk.

## A simple example Rmd file

Therefore a simple `.Rmd` file might be the following.

````{verbatim, lang="rmarkdown"}
---
title: "Example exercise"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE) # change to TRUE when knitting solutions
```

1. This is question 1. Which might have some R code you always want to show.
    ```{r, include=TRUE}
    # example code for the question
    ```
    
    ```{asis}
    Paragraph text for the solution can be kept in the document in an `asis` chunk. 
    And solution R code in an `r` chunk. 
    Both of these will use the `include` value from the `setup` chunk.
    ```
    
    ```{r}
    # example code for the solution
    ```

2. This is question 2.
````

To render the solutions output file we run the file with the following code.
````{verbatim, lang="rmarkdown"}
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```
````

## Parameterising our Rmd file

Instead of changing the `include` option in the setup chunk in the YAML header we can specify a `params` parameter with name of our choosing and pass it to `knitr::opts_chunk(include)` as follows.

````{verbatim, lang="rmarkdown"}
---
title: "Example exercise"
output: html_document
params:
  solutions: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = params$solutions)
```
````

## Customising the RStudio Knit button

The knitexercise package provides the `knit_exercise()` function to customise the Knit button.

You specify it in the YAML header as follows.
```yaml
---
title: "Example exercise"
output: html_document
params:
  solutions: FALSE
knit: knitexercise::knit_exercise
---
```

This function sets `knitr::opts_chunk(include)`. If knitting with the Knit button, when `solutions:` is set to `TRUE` it adds the suffix `-solutions` to the output filename before the file extension.

Customising the Knit button is described by Yihui Xie [here](https://bookdown.org/yihui/rmarkdown-cookbook/custom-knit.html).
