---
title: "Generating questions and solutions documents from the same R Markdown or Quarto document"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating questions and solutions documents from the same R Markdown or Quarto document}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This package provides a simple extra helper function when knitting an `.Rmd` file in which we have questions and solutions.

We ensure elements which are questions are either included as R Markdown text or from a code chunk with the chunk option `include=TRUE`.

We suppress or show the solutions with the knitr global document `include` chunk option controlled by a setup chunk at the beginning of the document as follows
````{verbatim}
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE) # set to TRUE for solutions document
```
````
Paragraph R Markdown text for the solutions is included in `asis` code chunks.

## A simple example Rmd file

Therefore a simple `.Rmd` file might be the following.

````{verbatim}
---
title: "Example exercise"
author: "My Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE) # change to TRUE when knitting solutions
```

1. This is question 1. Which might have some R code you always want to show.
   ```{r, include=TRUE}
   # example code for the question
   ```

   ```{asis}
   Paragraph text for the solution can be kept in the document in an `asis` chunk.
   And solution R code in an `r` chunk.
   Both of these will use the `include` value from the `setup` chunk.
   ```
    
   ```{r}
   # example code for the solution
   ```

2. This is question 2.
````

To render the solutions output file we run the file with the following code.
````{verbatim}
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```
````

Setting the global `include` chunk option also works in a Quarto document using the Knitr engine.

## Parameterising our Rmd file

Instead of changing the `include` option in the setup chunk in the YAML header we can specify a `params` parameter with name of our choosing and pass it to `knitr::opts_chunk(include)` as follows.

````{verbatim}
---
title: "Example exercise"
author: "My Name"
output: html_document
params:
  solutions: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = params$solutions)
```
<!-- ...Rest of document... -->
````

To generate both questions and solutions output documents from the single Rmd file we could run the following code at an R prompt.
```r
rmarkdown::render("exercise.Rmd", params = list(solutions = FALSE))
rmarkdown::render("exercise.Rmd", 
                  output_file = "exercise-solutions", 
                  params = list(solutions = TRUE))
```

### Parameterising a Quarto document

Quarto documents can also use the same parameterisation.
The parameters can be changed as arguments to the command
line interface or input in a `params.yml` file as described
[here](https://quarto.org/docs/computations/parameters.html).

```
For example, we may invoke the Quarto CLI as follows.
quarto render exercise.qmd -P solutions:TRUE
```
```
Or as follows if `params.yml` contains the line `solutions: TRUE`.
quarto render exercise.qmd --execute-params params.yml
```

Or rendering using the **quarto** R package we can change the parameters as follows.
```r
quarto::quarto_render("exercise.qmd", execute_params = list(solutions = "TRUE"))
quarto::quarto_render("exercise.qmd", execute_params = list(solutions = "FALSE"))
```

## Customising the RStudio Knit button

The **knitexercise** package provides the `knit_exercise()` function to customise the Knit button.

It is specified in the YAML header as follows.
```yaml
---
title: "Example exercise"
output: html_document
params:
  solutions: FALSE
knit: knitexercise::knit_exercise
---
```

This function sets `knitr::opts_chunk(include)`. If knitting with the Knit button, when `solutions:` is set to `TRUE` it adds the suffix `-solutions` to the output filename before the file extension.

Customising the Knit button is described by Yihui Xie [here](https://bookdown.org/yihui/rmarkdown-cookbook/custom-knit.html).
